The Near and Mid-infrared photometric properties of known redshift $z\geq5$ Quasars

### 2019-08-12 tex/MNRAS_submission/rv1/referees_reply.txt

p. 1: You may want to end the first sentence of the right column of the Intro
with a statement that at z>5, relativistic time dilation also stretches the
photometric variability of the accretion disk as observed.

To Do by NPR

p. 2, Fig. 1: Is it useful here to plot both filter sets if they are sufficiently
different? Please provide larger tick-marks for better readability in any case.

Hmmm. I’m not sure what to do here; it’s going to ruin the clarity of the figure
having both ZYJHK filter sets, but, they are different.
I think adding an explicit link to:
       d80b2t/VHzQ/SEDs/filters_vs_QSOstars_UKIRT.png
and
       d80b2t/VHzQ/SEDs/filters_vs_QSOstars_VISTA.png
is the way forward here.

p. 3, Fig. 3, caption or text: You should briefly comment on why the quasar
distribution has this gap in N(z) at z~5.5?. Is this because of the relativity
sensitivity across the filter curves, and/or the onset of complete GP-troughs at
z>5.7?

This is Figure 2, not 3.
And I think this is where some of the missing objects come in??
We’ll need to update: Tables 1, 2 and obviously 3 once we have those extra objects.

p. 4, Fig. 3: Can you plot the small dots somewhat bigger for better visibility
without loss of information?

This is the coverage map figure
       => To Do by NJGC.

THROUGHOUT: Several occurrences of ??, i.e., where table or Appendix references
are not properly done in latex.

NPR to double check…

p. 5: [My only (minor) scientific comment]: Re: ``although possible, we consider
it highly unlikely there would be more than one MIR source contributing to the
flux of a single UV/optically bright rest-frame quasar. Investigating this in
detail is very interesting but left to a future study. ''

I think it would be better to address now the number of expected contaminating
8--25 micron sources in the apertures used, as it can be done straightforwardly.
E.g., Driver et al. (2016, ApJ, 827, 107; astro-ph/1605.01523) --- the figures
you need for this estimate are actually best given as Fig. A4, A5 on the astro-ph
version. The redshift distribution of the sampled 8--25 micron foreground objects
is narrow enough and their K-corrections advantageous enough, that the objects
counts sharply turnover for AB(W3, W4) > 20--18 mag (left panels), and as a
consequence the normalized differential counts (right panels) --- especially in
W4 --- have most of their EBL energy emerging at AB~18--20 mag. As a consequence,
the relative number or probability of contaminating objects at AB>20 mag inside
your apertures should be rather small. A point worth mentioning, as it helps your
case.

NPR to add this to the text.

p. 6 Table 3: Should mention that the number of significant digits listed here
for the YJHK W1--W4 mags is not always warranted, but needed to verify the small
VIRCAM-WFCAM offsets in Table 5, This table is a very useful check of your global
photometric zeropoints.

NPR to change.

p. 7, very end of 3.1.1: Quote the (effective) VSA and SDSS quasar IR observing
epochs here, in case these clearly discrepant magnitudes are due to variability.

This is for SDSS J0349+0034 and J2220-0101; NJGC to check.

p. 8 Fig 5: Make the z=0--3 quasar cloud a little lighter grey, so that the
z=5--6.5 quasar cloud is better visible.

NPR to update.

p. 9 top left: = -0.458 : All minus signs in latex should be coded up as --
(double hyphen), also in the tables, to appear as minus sign.

NPR to update.

p. 10 Fig 8: Enlarge all axes label for better readability.

NPR to update.

THROUGHOUT: signal-to-noise should always be listed as a ratio: signal-to-noise
ratio or SNR.
NPR to update.

### 2019-08-19 tex/VHzQ.tex
wee updates

\subsection{Spectroscopy} 
We compile the list of all known, spectroscopically confirmed
quasars from the literature.  This list was complied from a range of
surveys and papers.  Specifically, we use data from:
\citet{Banados2014, Banados2016, Banados2018}, 
\citet{Becker2015}, 
\citet{Calura2014}, 
\citet{Carilli2007, Carilli2010}, 
\citet{Carnall2015}, 
\citet{Cool2006}, 
\citet{DeRosa2011}, 
\citet{Fan2000, Fan2001c, Fan2003, Fan2004, Fan2006, Fan2018}, 
\citet{Goto2006}, 
\citet{Ikeda2017}, 
\citet{Jiang2008, Jiang2009, Jiang2015, Jiang2016},   
\citet{Kashikawa2015}, 
\citet{Koptelova2017}, 
\citet{Kim2015, Kim2018},  
\citet{Kurk2007, Kurk2009}, 
\citet{Leipski2014}, 
\citet{Mahabal2005}, 
\citet{Matsuoka2016,  Matsuoka2018a, Matsuoka2018b},   
\citet{Mazzucchelli2017}, 
\citet{Morganson2012}, 
\citet{Mortlock2009, Mortlock2011},
\citet{McGreer2006, McGreer2013},  
\citet{Reed2015, Reed2017}, 
\citet{Stern2007},  
\citet{Tang2017}, 
\citet{Venemans2007, Venemans2012, Venemans2013, Venemans2015a, Venemans2015b, Venemans2016},
\citet{WangF2016, WangF2017, WangF2018a, WangF2018b},
\citet{Willott2007, Willott2009, Willott2010a, Willott2013b, Willott2015}, 
\citet{Wu2015} 
\citet{YangJ2018a, YangJ2018b}  
and 
\citet{Zeimann2011}.

@@

\citet{Secrest2015} present an all-sky sample of $\approx$1.35 million
AGN meeting a two-color, (W1-W2) vs. (W2-W3), photometric selection as
applied to sources from AllWISE. This colour-selection is motivated by
\citet{Mateos2012}. We investigate if this catalogue, which is tuned
for, and noted to be highly complete at, lower, $z\lesssim2$
redshifts, contain the VH$z$Qs. We find only 5 VH$z$Qs from our sample
match the Secrest catalogue, but given a W3 detection is required,
this is not surprising. Further, in addition to the relative lack of
sensitivity in W3, the mid-IR colors are moving out of the
\citet{Mateos2012} color criterion at high redshift. This was shown
using an AGN SED where AGNs above a redshift of about $\sim$2 fall out
of the colour criterion, see e.g. Figure 5 of \citet[][]{Mateos2012}
and Figure 9 of \citet{Secrest2015}.
%%
\citet{Assef2018} released two large catalogues of AGN
candidates identified across 30,000 deg$^2$ of extragalactic sky from
the WISE AllWISE Data Release. The ``R90'' catalogue, contains 4.5M
AGN candidates at 90\% reliability ($\approx$150 AGN candidates
per deg$^2$) while the ``C75'' catalogue consists of 20.9M AGN
candidates at 75\% completeness ($\approx$700 AGN candidates per
deg$^2$). Crossmatching our catalogue of 463 VH$z$Qs with these
catalogues, produces 42 matches with the R90 sample and 98 matches
with the C75 sample. Both catalogues unsurprisingly match to the
ultraluminous quasar SDSS J0100+2802 \citep{Wu2015} while the C75, but
not the R90 catalogue matches to the first redshift $z>7$ quasar ULAS
J1120+0641 \citep{Mortlock2011}. Neither the R90 or C75 catalogue matches to the
current highest-redshift object J1342+0928 \citep{Banados2018}.

\begin{figure*}
    \centering
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/DELS_J1048-0109.png} 
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/UHS_J0439+1634.png} 
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDWISE_J0306+1853.png}
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDSS_J0100+2802.png}
    \caption{VH$z$Qs with interesting MIR properties. 
      The thumbnails are the RGB colour outputs using the AllWISE W1W2W3 bands from the 
      IRSA WISE Image Service, with a scale of 120'' on the side. 
      {\it (a)} DELS J1048-0109 is the bluest in (W1-W2) colour quasar. 
      {\it (b)} UHS J0439+1634 was discovered by \citet{Fan2019} and is a strongly lensed quasar at $z=6.51$. 
      {\it (c)} SDWISE J0306+1853 has {\tt w3mpro} = $15.66\pm0.1$ (AB) and {\tt w4mpro } = $14.25\pm0.16$ (AB). 
      {\it (d)} SDSS J0100+2802 is the \citet{Wu2015} object.  
%      {\it (d)} SDSS J1443+3623 has {\tt w3mpro, w4mpro } = ($11.189\pm0.1$, $8.699\pm0.26$) and 
  %    {\it (d)} SDSS J1623+4705 has {\tt w3mpro, w4mpro } = ($11.425\pm0.1$, $8.954\pm0.28$).
}
    \label{fig:VHzQ_W4s}
\end{figure*}
\subsection{MIR Colours}
Due to the depth all-sky coverage of the WISE (and NEOWISE-(R))
mission, several investigations have looked at how WISE detects AGN
\citep[e.g][]{Stern2012, Assef2012, Secrest2015, LaMassa2017,
Assef2018, Glikman2018, Hviding2018, LaMassa2019}

Figure~\ref{fig:W1W2W3} shows the (W1-W2) vs. (W2-W3) colour-colour
diagram for the 283 VH$z$Qs that have reported {\tt w3mpro} values
(blue to red coloured points), though we note that only 51 (18\%) of
these objects have {\tt w3snr} $>$3.0. Approximately 400,000 quasars
at redshift $z\lesssim3.0$ from the SDSS Quasar Catalog using the
Fourteenth data release \citep[DR14Q; ][]{Paris2018} are also shown
(grey colour-scale).
% W1 samples $\approx$0.6-4$\mu$m for the $z=0-3$ quasars, 
% W2 samples $\approx$1.25-5$\mu$m for the $z=0-3$ quasars, 
% W3 samples $\approx$2.-16$\mu$m for the $z=0-3$ quasars,

@@

The set of VH$z$Q detected in WISE W3 and W4 contains 51 objects that are detected in the broad W3 filter and 10 objects that are formally detected in W4 ({\tt w4snr}$\geq$3.0). Three of the W3/4 objects are presented in Figure~\ref{fig:VHzQ_W4s}. UHS~J0439+1634 (Fig.~\ref{fig:VHzQ_W4s} {\it (b)}) with W4=7.165$\pm$0.12 was discovered by \citet{Fan2019} and is a strongly lensed quasar at $z=$6.51. This high luminosity is mostly not intrinsic, but is boosted by an intervening redshift $z\sim$0.7 galaxy. Other high-redshift quasars that are bright W3/4 objects e.g. SDWISE~J0306+1853 (Fig.~\ref{fig:VHzQ_W4s} {\it(c)}) may well also so be lensed \citep[e.g.,][]{Glikman2018lens, Fan2019, Pacucci2019}, but high-resolution follow-up is needed to confirm. SDSS~J0100+2802 (redshift $z=6.33$; Fig.~\ref{fig:VHzQ_W4s} {\it(d)}) does not have a formal W4 detection ({\tt w4snr}=2.4), but as reported by \citet{Wu2015}, has a detection in the 2MASS $J$, $H$ and $K_{\rm s}$ bands and is also the bluest object in both (W1-W3) and (W2-W3). %% with values of 2.73 (Vega) and 1.93 (Vega), respectively.

@@

\begin{figure}
  \includegraphics[width=8.5cm]
  {../light_curves/MIR_LCs/three_MIR_LC_egs_20180827.pdf}
  \centering
  \caption[]
  {Here we show the MIR NEOWISE-R for J0100+2802 \citep{Wu2015}, J0224-4711 and  J1626+2751. 
    Red points are the W1 band; cyan points the W2 band.} 
  \label{fig:MIR_LC_3egs}
\end{figure}
One key point to note when looking at the WISE variability is that
AllWISE (and 2MASS) source information included in the NEOWISE Source
Database
\href{http://wise2.ipac.caltech.edu/docs/release/neowise/expsup/sec2_1f.html}{are
associations not identifications}. The NEOWISE-AllWISE match radius is
3''.

Figure~\ref{fig:MIR_LC_epochs} gives the number of NEOWISE-R epochs and detections there are for each VH$z$Q.

while Figure~\ref{fig:MIR_LC_3egs} presents three examples of the MIR lightcurves and associated colour changes. 
Here we show J0100+2802 \citep{Wu2015}, J0224-4711 and J1626+2751.

@@

%\bibliographystyle{apj}
%\bibliographystyle{mn2e}
\bibliographystyle{mnras}
\bibliography{tester_mnras}

>>>>

\subsection{Spectroscopy} 
We compile the list of all known, spectroscopically confirmed
quasars from the literature.  This list was complied from a range of
surveys and papers.  Specifically, we use data from:
\citet{Banados2014, Banados2016, Banados2018a, Banados2018b}, 
\citet{Becker2015}, 
\citet{Calura2014}, 
\citet{Carilli2007, Carilli2010}, 
\citet{Carnall2015}, 
\citet{Cool2006}, 
\citet{Douglas2007}, 
\citet{DeRosa2011}, 
\citet{Fan2000, Fan2001c, Fan2003, Fan2004, Fan2006, Fan2018}, 
\citet{Goto2006}, 
\citet{Ikeda2017}, 
\citet{Jiang2008, Jiang2009, Jiang2015, Jiang2016},   
\citet{Kashikawa2015}, 
\citet{Khorunzhev2017}, 
\citet{Koptelova2017}, 
\citet{Kozlowski2019}, 
\citet{Kim2015, Kim2018},  
\citet{Kurk2007, Kurk2009}, 
\citet{Leipski2014}, 
\citet{Mahabal2005}, 
\citet{Matsuoka2016,  Matsuoka2018a, Matsuoka2018b},   
\citet{Mazzucchelli2017}, 
\citet{Morganson2012}, 
\citet{Mortlock2009, Mortlock2011},
\citet{McGreer2006, McGreer2013},  
\citet{Reed2015, Reed2017}, 
\citet{Saxena2018}, 
\citet{Stern2007},  
\citet{Tang2017}, 
\citet{vanBreugel1999}, 
\citet{Venemans2007, Venemans2012, Venemans2013, Venemans2015a, Venemans2015b, Venemans2016},
\citet{WangF2016, WangF2017, WangF2018a, WangF2018b},
\citet{Willott2007, Willott2009, Willott2010a, Willott2013b, Willott2015}, 
\citet{Wu2015} 
\citet{YangJ2018a, YangJ2018b}  
and 
\citet{Zeimann2011}.

@@

\citet{Secrest2015} present an all-sky sample of $\approx$1.35 million
AGN meeting a two-color, (W1-W2) vs. (W2-W3), photometric selection as
applied to sources from AllWISE. This colour-selection is motivated by
\citet{Mateos2012}. We investigate if this catalogue, which is tuned
for, and noted to be highly complete at, lower, $z\lesssim2$
redshifts, contain the VH$z$Qs. We find only 5 VH$z$Qs from our sample
match the Secrest catalogue, but given a W3 detection is required,
this is not surprising. Further, in addition to the relative lack of
sensitivity in W3, the mid-IR colors are moving out of the
\citet{Mateos2012} color criterion at high redshift. This was shown
using an AGN SED where AGNs above a redshift of about $\sim$2 fall out
of the colour criterion, see e.g. Figure 5 of \citet[][]{Mateos2012}
and Figure 9 of \citet{Secrest2015}.
%%
\citet{Assef2018} released two large catalogues of AGN
candidates identified across 30,000 deg$^2$ of extragalactic sky from
the WISE AllWISE Data Release. The ``R90'' catalogue, contains 4.5M
AGN candidates at 90\% reliability ($\approx$150 AGN candidates
per deg$^2$) while the ``C75'' catalogue consists of 20.9M AGN
candidates at 75\% completeness ($\approx$700 AGN candidates per
deg$^2$). Crossmatching our catalogue of 463 VH$z$Qs with these
catalogues, produces 42 matches with the R90 sample and 98 matches
with the C75 sample. Both catalogues unsurprisingly match to the
ultraluminous quasar SDSS J0100+2802 \citep{Wu2015} while the C75, but
not the R90 catalogue matches to the first redshift $z>7$ quasar ULAS
J1120+0641 \citep{Mortlock2011}. Neither the R90 or C75 catalogue matches to the
current highest-redshift object J1342+0928 \citep{Banados2018a}.

\begin{figure*}
    \centering
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/DELS_J1048-0109.png} 
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/UHS_J0439+1634.png} 
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDWISE_J0306+1853.png}
    \includegraphics[width=0.24\textwidth]{../data/WISE/image_thumbnails/SDSS_J0100+2802.png}
    \caption{VH$z$Qs with interesting MIR properties. 
      The thumbnails are the RGB colour outputs using the AllWISE W1W2W3 bands from the 
      IRSA WISE Image Service, with a scale of 120'' on the side. 
      {\it (a)} DELS J1048-0109 is the bluest in (W1-W2) colour quasar. 
      {\it (b)} UHS J0439+1634 was discovered by \citet{Fan2019} and is a strongly lensed quasar at $z=6.51$. 
      {\it (c)} SDWISE J0306+1853 has {\tt w3mpro} = $15.66\pm0.1$ (AB) and {\tt w4mpro } = $14.25\pm0.16$ (AB). 
      {\it (d)} SDSS J0100+2802 is the \citet{Wu2015} object.}
    \label{fig:VHzQ_W4s}
\end{figure*}
\subsection{MIR Colours}
Due to the depth all-sky coverage of the WISE (and NEOWISE-(R))
mission, several investigations have looked at how WISE detects AGN
\citep[e.g][]{Stern2012, Assef2012, Secrest2015, LaMassa2017,
Assef2018, Glikman2018, Hviding2018, LaMassa2019}

Figure~\ref{fig:W1W2W3} shows the (W1-W2) vs. (W2-W3) colour-colour
diagram for the 283 VH$z$Qs that have reported {\tt w3mpro} values
(blue to red coloured points), though we note that only 51 (18\%) of
these objects have {\tt w3snr} $>$3.0. Approximately 400,000 quasars
at redshift $z\lesssim3.0$ from the SDSS Quasar Catalog using the
Fourteenth data release \citep[DR14Q; ][]{Paris2018} are also shown
(grey colour-scale).

@@

The set of VH$z$Q detected in WISE W3 and W4 contains 51 objects that
are detected in the broad W3 filter and 10 objects that are formally
detected in W4 ({\tt w4snr}$\geq$3.0). Three of the W3/4 objects are
presented in Figure~\ref{fig:VHzQ_W4s}. UHS~J0439+1634
(Fig.~\ref{fig:VHzQ_W4s} {\it (b)}) with W4=7.165$\pm$0.12 was
discovered by \citet{Fan2019} and is a strongly lensed quasar at
$z=$6.51. This high luminosity is mostly not intrinsic, but is boosted
by an intervening redshift $z\sim$0.7 galaxy. Other high-redshift
quasars that are bright W3/4 objects e.g. SDWISE~J0306+1853
(Fig.~\ref{fig:VHzQ_W4s} {\it(c)}) may well also so be lensed
\citep[e.g.,][]{Glikman2018lens, Fan2019, Pacucci2019}, but
high-resolution follow-up is needed to confirm. SDSS~J0100+2802
(redshift $z=6.33$; Fig.~\ref{fig:VHzQ_W4s} {\it(d)}) does not have a
formal W4 detection ({\tt w4snr}=2.4), but as reported by
\citet{Wu2015}, has a detection in the 2MASS $J$, $H$ and $K_{\rm s}$
bands and is also the bluest object in both (W1-W3) and (W2-W3).

@@

\begin{figure}
  \centering
  \includegraphics[width=8.5cm]{../light_curves/MIR_LCs/NEOWISER_LCs_VHzQs_SDSS_J0100.png}
  \includegraphics[width=8.5cm]{../light_curves/MIR_LCs/NEOWISER_LCs_VHzQs_SDWISE_J1752.png}
  \includegraphics[width=8.5cm]{../light_curves/MIR_LCs/NEOWISER_LCs_VHzQs_SDSS_J1621.png}
  \caption[]
  {Here we show the MIR NEOWISE-R for SDSS J0100+2802 (top), SDWISE   J1752+5036 (middle), and  
    SDSS J1621+5155 (bottom). Red points are the W1 band; cyan points the W2 band.
  For the light curves, we show weighted means for every 6 month NEOWISE-R scan.} 
  \label{fig:MIR_LC_3egs}
\end{figure}
One key point to note is that the NEOWISE-AllWISE match radius 
is 3.00'' and the AllWISE (and 2MASS) source information included 
in the NEOWISE Source Database 
\href{http://wise2.ipac.caltech.edu/docs/release/neowise/expsup/sec2_1f.html}{are
associations not identifications}.

Figure~\ref{fig:MIR_LC_epochs} gives the number of NEOWISE-R epochs and detections for each VH$z$Q. 
There are 60 quasars with over 100 NEOWISE-R epochs, nine with over 200 epochs and four with over 300.
Figure~\ref{fig:MIR_LC_3egs} presents three examples of the brighter MIR objects with well-sampled 
MIR lightcurves and associated colour changes; 
SDSS J0100+2802 (Fig.~\ref{fig:MIR_LC_3egs} top), 
SDWISE  J1752+5036    (middle), 
SDSS J1621+5155 (bottom). 
For the light curves, we show weighted means for every 6 month NEOWISE-R scan.
Due to the WISE scanning pattern, objects close to the North Ecliptic Pole 
are well sampled, and indeed this is the area for further investigation 
time-domain investigation with upcoming large infrared space telescopes
\citep{Jansen2018, White2019AAS}.

@@

%\bibliographystyle{apj}
%\bibliographystyle{mn2e}
\bibliographystyle{mnras}
\bibliography{/cos_pc19a_npr/programs/quasars/highest_z/tex/tester_mnras}
